Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-16 special-issue-ci/Response to reviewers.tex

\begin{quote}
-  an experiment with a more difficult problem (please use the well-known benchmark problems
described in literature).
\end{quote}

A experiment using the well known Rastrigin function has been
performed and analyzed.

\begin{quote}
-  A comparison in similar conditions of the two implementations of the architecture (the algorithms
used in the experiments are slightly different, but this could lead to dramatic changes in
performance, and consequently, invalidate the comparison performed in the paper). This would
enable a measurement of the actual improvement in performance obtained by the use of web
workers (not that generated due to a different algorithm)
\end{quote}

This has been included as the last subsection of the paper. The new
experiment has constant population, instead of using random
population.

@@

\begin{quote}
Another issue is that some details are described twice, for instance, the advantages of the use of web
workers. Fixing this issue can reduce the paper size and improve its
readability.
\end{quote}

We have eliminated duplications in this new version. Thanks for
noticing it.

\begin{quote}
Regarding the presentation:
-  Figures should be readable even in gray-scale, so do not use red and blue to denote different
elements, use a different shade or texture.
-  The axis labels and legends in most of the figures are unreadable, please increase its font size.
-  There are some typos and phrases that need rewriting.
\end{quote}

We have rewritten and revised extensively the paper, and tried to find
a way (and failed) to make it more readable for reviewers. In fact,
they use different shades, to it should mostly work in grayscale. For the time being and
during the revision phase, we prefer to keep it that way, but we will
change it to grayscale if it is required for the final version.

### 2015-11-16 special-issue-ci/Response to reviewers.tex

\begin{quote}
Another issue is that some details are described twice, for instance, the advantages of the use of web
workers. Fixing this issue can reduce the paper size and improve its
readability.
\end{quote}

We have eliminated duplications in this new version. Thanks for
noticing it.

\begin{quote}
Regarding the presentation:
-  Figures should be readable even in gray-scale, so do not use red and blue to denote different
elements, use a different shade or texture.
-  The axis labels and legends in most of the figures are unreadable, please increase its font size.
-  There are some typos and phrases that need rewriting.
\end{quote}

We have rewritten and revised extensively the paper, and tried to find
a way (and failed) to make it more readable for reviewers. In fact,
they use different shades, to it should mostly work in grayscale. For the time being and
during the revision phase, we prefer to keep it that way, but we will
change it to grayscale if it is required for the final version. 
In fact, the labels are not so important, so increasing the font size
would in fact distract the attention from the figure itself, which is
the interesting information we want to convey with the graph. Once
again, we can do it if required for the final version.

@@

\section{Response to Reviewer \#3}

\begin{quote}
1. Contribution: The paper presents the idea of a web browser based distributed evolutionary computing.
the system called NodIO consists of a REST server (coded in Node.js) and clients performing
evolutionary algorithm coded in JavaScript and embedded in a web page. Although the idea is not
entirely new, the authors show how Web Workers (introduced in HTML5) and the mechanism of
automatic restart of computations on the client's side can improve the performance of the system. The
state of the art is comprehensive and provides good introduction into the subject.

2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
Software" special issue. Thanks to both development of Internet technology and continuous increase in
computing power of electronic devices, including mobile devices, the web browser based distributed
computing can be a cheap alternative for grid and high-performance computing systems that require
high capital expenditures and consume a lot of energy. Evolutionary computing system based on the
architecture presented inthe paper can be potentially used for solving many different time consuming
problems, however, the authors limited their experiments to a simple "Trap" function that can be easily
solved on a desktop machine, what may not be convincing for a reader looking for an effective
optimization tool.
\end{quote}

It is true that that function can be easily solved with any desktop
machine, but we can still do that {\em and} use a volunteer
computation framework, which will add performance to the desktop
machine. Our intention was more to prove what kind of performance can
be expected and how this performance is distributed, although we have
also proved in the last section how the performance achieved by a
volunteer computer system can, in fact, be better than a desktop, even
a powerful one.

At any rate, we have also performed a new experiment with the
Rasrtrigin function, which in fact cannot be solved in a few minutes
(or seconds, as in the case of Trap) by a desktop system.

\begin{quote}
3. Correctness: The system works correctly - it is not very complex and the authors provide its source
code (this is a great plus!). Technical description of the system is also clear, however it lacks some
details like parameters of the machine used to determine baselines for the experiments or the format of
messages sent between the server and the clients. The only thing I noticed is that the authors used the
term "multiple CPUs" in the description of Web Worker API advantages (page 8) instead of "multi-core
CPUs". As far as I know, contemporary web browsers do not support multiple CPUs architecture.

### 2015-11-16 special-issue-ci/Response to reviewers.tex

\section{Response to Reviewer \#3}

\begin{quote}
1. Contribution: The paper presents the idea of a web browser based distributed evolutionary computing.
the system called NodIO consists of a REST server (coded in Node.js) and clients performing
evolutionary algorithm coded in JavaScript and embedded in a web page. Although the idea is not
entirely new, the authors show how Web Workers (introduced in HTML5) and the mechanism of
automatic restart of computations on the client's side can improve the performance of the system. The
state of the art is comprehensive and provides good introduction into the subject.

2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
Software" special issue. Thanks to both development of Internet technology and continuous increase in
computing power of electronic devices, including mobile devices, the web browser based distributed
computing can be a cheap alternative for grid and high-performance computing systems that require
high capital expenditures and consume a lot of energy. Evolutionary computing system based on the
architecture presented inthe paper can be potentially used for solving many different time consuming
problems, however, the authors limited their experiments to a simple "Trap" function that can be easily
solved on a desktop machine, what may not be convincing for a reader looking for an effective
optimization tool.
\end{quote}

It is true that that function can be easily solved with any desktop
machine, but we can still do that {\em and} use a volunteer
computation framework, which will add performance to the desktop
machine. Our intention was more to prove what kind of performance can
be expected and how this performance is distributed, although we have
also proved in the last section how the performance achieved by a
volunteer computer system can, in fact, be better than a desktop, even
a powerful one. This has also been clarified in the introduction and
conclusions.

@@

At any rate, we have also performed a new experiment with the
Rasrtrigin function, which in fact cannot be solved in a few minutes
(or seconds, as in the case of Trap) by a desktop system.

\begin{quote}
3. Correctness: The system works correctly - it is not very complex and the authors provide its source
code (this is a great plus!). Technical description of the system is also clear, however it lacks some
details like parameters of the machine used to determine baselines for the experiments or the format of
messages sent between the server and the clients. The only thing I noticed is that the authors used the
term "multiple CPUs" in the description of Web Worker API advantages (page 8) instead of "multi-core
CPUs". As far as I know, contemporary web browsers do not support
multiple CPUs architecture.
\end{quote}

This has been clarified in the new version of the paper, third
paragraph from the beginning of the section that explains the web
worker-based architecture. In fact, most browsers can use multi-core
CPUs running web workers as independent processes in each one of them.
The base system has been now described in the first paragraph of the
section on the experiments, it a desktop Intel i7-4770.

\begin{quote}

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Details on baseline computer added.
Closes #42

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{baseline-times.png}
\caption{Comparison of time to solution for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using NodEO
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with success rate upgraded to 100\% and average duration $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}.

@@

Initial experiments were set up using the OpenShift
PaaS, which provides a free tier within which this 
experiment could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the months of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. No special care was taken to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is, in fact negligible.
%
\begin{table*}
\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rr}
\hline
Date & \# experiments & Different IPs \\
\hline
April 4th 4/4 & 57 & 191 \\
April 24th 4/24 &  231 & 559 \\
July 31th 7/31 & 97 & 179 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
\begin{table*}
\caption{Summary of time per run, number of IPS and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
\hline
\end{tabular}
\end{center}
\end{table*}
%
The table shows that every run included more than 50 experiments. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment.

>>>>

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{baseline-times.png}
\caption{Comparison of time to solution for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using NodEO
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with success rate upgraded to 100\% and average duration $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline system has these
characteristics: \begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with an 8-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

@@

Initial experiments were set up using the OpenShift
PaaS, which provide a free tier within which this 
experiment could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the months of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. No special care was taken to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is, in fact negligible.
%
\begin{table*}
\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rr}
\hline
Date & \# experiments & Different IPs \\
\hline
April 4th 4/4 & 57 & 191 \\
April 24th 4/24 &  231 & 559 \\
July 31th 7/31 & 97 & 179 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
\begin{table*}
\caption{Summary of time per run, number of IPS and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
\hline
\end{tabular}
\end{center}
\end{table*}
%
The table shows that every run included more than 50 experiments. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment.
